---
title: "4. Plot Microbiome Network"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 4. Plot Microbiome Network

```{r}
## Load Package and Example Data
library(phyloseq); packageVersion("phyloseq")
packageVersion("ggplot2")
data(enterotype)
```

```{r}
## set the random number generator seed explicitly
set.seed(711L)
```

#### this code is to remove the 9 samples for which no enterotype designation was assigned, but it was unable to knit. Thus 9 samples are still included below plots.
    enterotype = subset_samples(enterotype, !is.na(Enterotype))

## The plot_net function

```{r}
## The default setting
plot_net(enterotype, maxdist = 0.4, point_label = "Sample_ID")
```

```{r}
plot_net(enterotype, maxdist = 0.3, color = "SeqTech", shape="Enterotype")
```

## The plot_network function
Create an igraph-based network based on the default distance method, “Jaccard”, and a maximum distance between connected nodes of 0.3.
```{r}
ig <- make_network(enterotype, max.dist=0.3)
plot_network(ig, enterotype)
```

```{r}
plot_network(ig, enterotype, color="SeqTech", shape="Enterotype", line_weight=0.4, label=NULL)
```

```{r}
## when the maximum distance is lowered, decreasing the number of edges in the network
ig <- make_network(enterotype, max.dist=0.2)
plot_network(ig, enterotype, color="SeqTech", shape="Enterotype", line_weight=0.4, label=NULL)
```

```{r}
## replace the Jaccard (default) distance method with Bray-Curtis
ig <- make_network(enterotype, dist.fun="bray", max.dist=0.3)
plot_network(ig, enterotype, color="SeqTech", shape="Enterotype", line_weight=0.4, label=NULL)
```
