---
title: "Lab3 a: Starting with the Data"
author: "Hyejoo Kang"
date: "9/7/2020"
output:
  rmarkdown::html_document:
    theme: lumen
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

How to load you data file from external database into a data frame? Understand what data frame is and how to inspect data using R and RStudio. To see the original tutorial [click here](https://datacarpentry.org/R-ecology-lesson/02-starting-with-data.html).

# Presentation of the Survey Data
* Loading tidyverse

```{r}
library(tidyverse)
```


* Downloading cvs file
```{r, eval = FALSE}
download.file(url = "https://ndownloader.figshare.com/files/2292169", destfile = "data/portal_data_joined.csv")
```

* Loading the data
```{r}
surveys <- read.csv("data/portal_data_joined.csv")
```
* View the top 6 lines of the data frame
```{r}
head(surveys)
```
* View the data in new tap
```{r, eval = FALSE}
View(surveys)
```

# Learn about data frame
```{r}
str(surveys)
```

# Inspecting 'data.frame' Objects
* Size:
```{r, eval = FALSE}
dim(surveys)
nrow(surveys)
ncol(surveys)
```
* Content:
```{r, eval = FALSE}
head(surveys)
tail(surveys)
```
* Names:
```{r, eval = FALSE}
names(surveys)
rownames(surveys)
```
* Summary:
```{r, eval = FALSE}
summary(surveys)
```

## Challenge 1
*Based on the output of str(surveys), can you answer the following questions?*

* What is the class of the object surveys?
```{r}
class(surveys)
```

* How many rows and how many columns are in this object?
```{r}
nrow(surveys)
```
```{r}
ncol(surveys)
```
* How many species have been recorded during these surveys?
```{r}
length(unique(surveys$species))
```
```{r}
length(unique(surveys$species_id))
```

# Indexing and subsetting data frames
Run 'survey[row#, column#]' to index some specific data as a vector. Running 'surveys[column#]' or 'surveys [row#, ]' display data as a data.frame.

* 3rd column of the data frame (as a vector)
```{r, eval = FALSE}
surveys[, 3]
```

* 2nd column of the data frame (as a data.frame)
```{r, eval = FALSE}
surveys[2]
```

':' is a special function for numeric vectors of integers 

* first two elements in the 10th column (as a vector)
```{r, eval = FALSE}
surveys[1:2, 10]
```

Data frames and vectors can be subset by direct name of column
```{r, eval = FALSE}
surveys["species_id"]
surveys[, "species_id"]
```

## Challenge 2
*1. Create a data.frame(surveys_200) containing only the data in row 200 of the survey dataset*
```{r}
surveys_200 <- surveys[200, ]
```

*2. Notice how 'nrow()' gave you the number of rows in a data.frame?*
```{r}
nrow(surveys)
nrow(surveys_200)
```
* Use that number to pull out just that last row in the data frame.
```{r}
surveys[34786, ]
surveys_200[1, ]
```

* Compare that with what you see as the last row using 'tail()' to make sure it’s meeting expectations.
```{r}
tail(surveys)
tail(surveys_200)
```

* Pull out that last row using 'nrow()' instead of the row number.
```{r}
n_rows <- nrow(surveys)
```

* Create a new data frame (surveys_last) from that last row.
```{r}
surveys_last <- surveys[n_rows, ]
```

*3. Use 'nrow()' to extract the row that is in the middle of the data frame. Store the content of this row in an object named 'surveys_middle'.*

```{r}
n_rows / 2
```

```{r}
surveys_middle <- surveys[17393/2, ]
surveys_middle
```

*4. Combine 'nrow()' with the - notation above to reproduce the behavior of 'head(surveys)', keeping just the first through 6th rows of the surveys dataset.*
```{r}
surveys[-(7:n_rows), ]
survyes <- surveys[-(7:n_rows), ]
```

# Factors
Factors represent categorical data. Factors contains only a pre-defined levels and levels can be inspected...
```{r}
sex <- factor(c("male", "female", "female", "male"))
levels(sex)
nlevels(sex)
```

You can also specify the order of levels for you analysis...

```{r}
sex # current order
```

```{r}
sex <- factor(sex, levels = c("male", "female"))
sex # after re-ordering
```

## Converting factors

```{r}
as.character(sex)
```
## Renaming factors
```{r}
plot(as.factor(surveys$sex))
```
```{r}
sex_surveys <- factor(surveys$sex)
head(sex_surveys)
```
```{r}
levels(sex_surveys)
```
```{r}
levels(sex_surveys)[1] <- "undetermined"
```

```{r}
levels(sex_surveys)
```

```{r}
head(sex_surveys)
```

## Challenge 3
* Rename “F” and “M” to “female” and “male” respectively.
```{r}
levels(sex_surveys)[2:3] <- c("female", "male")
```

* Now that we have renamed the factor level to “undetermined”, can you recreate the barplot such that “undetermined” is last (after “male”)?
```{r}
sex <- factor(sex_surveys, levels = c("female", "male", "undetermined"))
```
```{r}
plot(sex)
```

## 'stringsAsFactors = FALSE'
When our data read columns' characters (strings) as 'factor' vs 'character'
```{r, eval=FALSE}
surveys <- read.csv("data/portal_data_joined.csv", stringsAsFactors = TRUE)
surveys
```

```{r, eval=FALSE}
surveys <- read.csv("data/portal_data_joined.csv", stringsAsFactors = FALSE)
str(surveys)
```
Converting the column "plot_type" into a factor
```{r}
surveys$plot_type <- factor(surveys$plot_type)
```

## Challenge 4
*1. Fix this hand-crafted 'data.frame()'!*
```{r}
animal_data <- data.frame(
          animal = c("dog", "cat", "sea cucumber", "sea urchin"),
          feel = c("furry", "furry", "squishy", "spiny"),
          weight = c(45, 8, 1.1, 0.8)
          )
```

*2. Can you predict the class for each of the columns in the following example? Check your guesses using str(country_climate):*
```{r}
country_climate <- data.frame(
       country = c("Canada", "Panama", "South Africa", "Australia"),
       climate = c("cold", "hot", "temperate", "hot/temperate"),
       temperature = c(10, 30, 18, "15"),
       northern_hemisphere = c(TRUE, TRUE, FALSE, "FALSE"),
       has_kangaroo = c(FALSE, FALSE, FALSE, 1)
       )
```
* Are they what you expected? Why? Why not?
```{r}
str(country_climate)
```
* What would have been different if we had added 'stringsAsFactors = FALSE' when creating the data frame:
'stringsAsFactors = FALSE' will convert strings to just characters, not factors.

* What would you need to change to ensure that each column had the accurate data type?
```{r}
country_climate_fixed <- data.frame(
       country = c("Canada", "Panama", "South Africa", "Australia"),
       climate = c("cold", "hot", "temperate", "hot/temperate"),
       temperature = c(10, 30, 18, 15),
       northern_hemisphere = c(TRUE, TRUE, FALSE, FALSE),
       has_kangaroo = c(FALSE, FALSE, FALSE, TRUE)
       )
```
```{r}
str(country_climate_fixed)

```

# Formatting Dates
```{r}
library(lubridate)
```

```{r}
my_date <- ymd("2020-01-01")
str(my_date)
```

```{r}
my_date <- ymd(paste("2020", "1", "1", sep = "-")) 
str(my_date)
```

```{r, eval=FALSE}
paste(surveys$year, surveys$month, surveys$day, sep = "-")
```

```{r, eval=FALSE}
ymd(paste(surveys$year, surveys$month, surveys$day, sep = "-"))
```

```{r, eval=FALSE}
surveys$date <- ymd(paste(surveys$year, surveys$month, surveys$day, sep = "-"))
```

```{r}
str(surveys)
```

```{r}
summary(surveys$date)
```

```{r}
missing_dates <- surveys[is.na(surveys$date), c("year", "month", "day")]
head(missing_dates)
```

