---
title: "Lab4 Working with COVID-19 reporting data"
author: "Hyejoo Kang"
date: "9/14/2020"
output:
  rmarkdown::html_document:
    theme: lumen
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(DT)
```

# Covid 19 daily report on March 13, 2020
```{r}
report_03_13_2020 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-13-2020.csv")) %>% 
  rename(Country_Region = "Country/Region", Province_State = "Province/State")
```

```{r}
str(report_03_13_2020)
```

```{r, eval=FALSE}
View(report_03_13_2020)
```
### Data table, March 13, 2020
```{r}
datatable(report_03_13_2020)
```

# Covid 19 daily report on September 13, 2020
```{r}
report_09_13_2020 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/09-13-2020.csv"))
```

```{r}
str(report_09_13_2020)
```
```{r, eval=FALSE}
view(report_09_13_2020)
```
### Data table, September 13, 2020
```{r}
datatable(report_09_13_2020)
```

# March vs September: Total counts by country
### March 13, 2020 
```{r}
report_march_total <- report_03_13_2020 %>% 
  group_by(Country_Region) %>% 
  summarize(Total_Confirmed = sum(Confirmed),
            Total_Deaths = sum(Deaths),
            Total_Recovered = sum(Recovered))
datatable(report_march_total)


```

### September 13, 2020
```{r}
report_september_total <- report_09_13_2020 %>% 
  group_by(Country_Region) %>% 
  summarize(Total_Confirmed = sum(Confirmed),
            Total_Deaths = sum(Deaths),
            Total_Recovered = sum(Recovered))
datatable(report_september_total)
```


# March vs September: Total counts by US state

### March 13, 2020
```{r}
report_march_US_total <- report_03_13_2020 %>% 
  filter(Country_Region == "US") %>% 
  group_by(Province_State) %>% 
  summarize(Total_Confirmed = sum(Confirmed),
            Total_Deaths = sum(Deaths),
            Total_Recovered = sum(Recovered))
datatable(report_march_US_total)
```

### September 13, 2020
```{r}
report_september_US_total <- report_09_13_2020 %>% 
  filter(Country_Region == "US") %>% 
  filter(Province_State != "Recovered") %>% 
  group_by(Province_State) %>% 
  summarize(Total_Confirmed = sum(Confirmed),
            Total_Deaths = sum(Deaths),
            Total_Recovered = sum(Recovered))
datatable(report_september_US_total)
```


# March Vs September: Top 10 countries for Deaths

### March 13, 2020
```{r}
report_march_top_10 <- report_03_13_2020 %>% 
  group_by(Country_Region) %>% 
  summarize(Total_Confirmed = sum(Confirmed),
            Total_Deaths = sum(Deaths),
            Total_Recovered = sum(Recovered)) %>% 
  arrange(desc(Total_Deaths)) %>% 
  slice(1:10)
datatable(report_march_top_10)
```

### September 13, 2020
```{r}
report_sept_top_10 <- report_09_13_2020 %>% 
  group_by(Country_Region) %>% 
  summarize(Total_Confirmed = sum(Confirmed),
            Total_Deaths = sum(Deaths),
            Total_Recovered = sum(Recovered)) %>% 
  arrange(desc(Total_Deaths)) %>% 
  slice(1:10)
datatable(report_sept_top_10)
```

# Plot of top 10 countries with the Confirmed case & Deaths (geom_point)

```{r}
library(scales)
```

```{r}
Thousand <- function (x) { number_format(accuracy = 1,
                                   scale = 1/1000,
                                   suffix = "k",
                                   big.mark = ",")(x) }
```

### Confirmed cases on March 13, 2020
```{r}
ggplot(report_march_top_10, aes(x = Total_Confirmed, y = Country_Region)) + 
  geom_point() + 
  scale_x_continuous(name = "Confirmed", label = Thousand)
```

### Confirmed cases on September 13, 2020
```{r}
ggplot(report_sept_top_10, aes(x = Total_Confirmed, y = Country_Region)) + 
  geom_point() + 
  scale_x_continuous(name = "Confirmed", label = Thousand)
```

### Deaths on March 13, 2020
```{r}
ggplot(report_march_top_10, aes(x = Total_Deaths, y = Country_Region)) + 
  geom_point() + 
  scale_x_continuous(name = "Deaths", label = Thousand)
```

### Deaths on September 13, 2020
```{r}
ggplot(report_sept_top_10, aes(x = Total_Deaths, y = Country_Region)) + 
  geom_point() + 
  scale_x_continuous(name = "Deaths", label = Thousand)
```

# Plot of top 10 countries with the Confirmed case & Deaths (geom_bar)
```{r}
library(RColorBrewer)
```

### Confirmed cases on March 13, 2020
```{r}
  ggplot(report_march_top_10, aes(x = Total_Confirmed,  y = Country_Region, fill = Country_Region)) + 
  geom_bar(stat = 'identity') + 
  scale_x_continuous(name = "Confirmed", label = Thousand) +
  theme(legend.position="none")
```


### Confirmed cases on September 13, 2020
```{r}
ggplot(report_sept_top_10, aes(x = Total_Confirmed, y = Country_Region, fill = Country_Region)) + 
  geom_bar(stat = 'identity') + 
  scale_x_continuous(name = "Confirmed", label = Thousand) +
  theme(legend.position="none")
```

### Deaths on March 13, 2020
```{r}
ggplot(report_march_top_10, aes(x = Total_Deaths, y = Country_Region, fill = Country_Region)) + 
         geom_bar(stat = 'identity') + 
         scale_x_continuous(name = "Deaths", label = Thousand) +
  theme(legend.position="none")
```

### Deaths on September 13, 2020
```{r}
ggplot(report_sept_top_10, aes(x = Total_Deaths, y = Country_Region, fill = Country_Region)) + 
  geom_bar(stat = 'identity') + 
  scale_x_continuous(name = "Deaths", label = Thousand) +
  theme(legend.position="none")
```
