---
title: "Lab4a: Data visualization with ggplot2"
author: "Hyejoo Kang"
date: "9/14/2020"
output:
  rmarkdown::html_document:
    theme: lumen
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
```

```{r}
surveys_complete <- read_csv("data/surveys_complete.csv")
```

# Plotting with ggplot2

Use the ggplot() function and define an aesthetic mapping using aes() function
```{r}
ggplot(data = surveys_complete, mapping = aes(x = weight, y = hindfoot_length))
```

  Add 'geoms' for graphical representations of the data in the plot
* geom_point() for scatter plots, dot plots, etc.
* geom_boxplot() for, well, boxplots!
* geom_line() for trend lines, time series, etc.

```{r}
ggplot(data = surveys_complete, aes(x = weight, y = hindfoot_length)) +
  geom_point()
```


  The + sign used to add layers must be placed at the end of each line containing a layer. If, instead, the + sign is added in the line before the other layer, ggplot2 will not add the new layer and will return an error message.

```{r}
# Assign plot to a variable
surveys_plot <- ggplot(data = surveys_complete, 
                       mapping = aes(x = weight, y = hindfoot_length))

# Draw the plot
surveys_plot + 
    geom_point()
```

## Challenge
Scatter plots can be useful exploratory tools for small datasets. For data sets with large numbers of observations, such as the surveys_complete data set, overplotting of points can be a limitation of scatter plots. One strategy for handling such settings is to use hexagonal binning of observations. The plot space is tessellated into hexagons. Each hexagon is assigned a color based on the number of observations that fall within its boundaries. To use hexagonal binning with ggplot2, first install the R package hexbin from CRAN:

```{r}
library("hexbin")
```

```{r}
surveys_plot + 
  geom_hex()
```

# Building your plots iteratively

Add transparency (alpha) to avoid overplotting
```{r}
ggplot(data = surveys_complete, aes(x = weight, y = hindfoot_length)) +
    geom_point(alpha = 0.1)
```

  Add colors for all the points
```{r}
ggplot(data = surveys_complete, mapping = aes(x = weight, y = hindfoot_length)) +
    geom_point(alpha = 0.1, color = "blue")
```
```{r}
ggplot(data = surveys_complete, mapping = aes(x = weight, y = hindfoot_length)) +
    geom_point(alpha = 0.1, aes(color = species_id))
```

## Challenge
Use what you just learned to create a scatter plot of weight over species_id with the plot types showing in different colors. Is this a good way to show this type of data?

```{r}
ggplot(data = surveys_complete, mapping = aes(x = species_id, y = weight)) + 
  geom_point(aes(color = plot_type))
```

# Boxplot
```{r}
ggplot(data = surveys_complete, mapping = aes(x = species_id, y = weight)) +
    geom_boxplot()
```

```{r}
ggplot(data = surveys_complete, mapping = aes(x = species_id, y = weight)) +
    geom_boxplot(alpha = 0) +
    geom_jitter(alpha = 0.3, color = "tomato")
```

```{r}
ggplot(data = surveys_complete, mapping = aes(x = species_id, y = weight)) + 
  geom_jitter(alpha = 0.3, color = "tomato") + 
  geom_boxplot(alpha = 0) 
```

## Challenge
Boxplots are useful summaries, but hide the shape of the distribution. For example, if there is a bimodal distribution, it would not be observed with a boxplot. An alternative to the boxplot is the violin plot (sometimes known as a beanplot), where the shape (of the density of points) is drawn.
* Replace the box plot with a violin plot; see geom_violin().

```{r}
ggplot(data = surveys_complete, mapping = aes(x = species_id, y = weight)) + 
  geom_violin(alpha = 0) 
```

In many types of data, it is important to consider the scale of the observations. For example, it may be worth changing the scale of the axis to better distribute the observations in the space of the plot. Changing the scale of the axes is done similarly to adding/modifying other components (i.e., by incrementally adding commands). Try making these modifications:
* Represent weight on the log10 scale; see scale_y_log10().

```{r}
ggplot(data = surveys_complete, mapping = aes(x = species_id, y = weight)) + 
  geom_violin(alpha = 0) + 
  scale_y_log10()
```

So far, we’ve looked at the distribution of weight within species. Try making a new plot to explore the distribution of another variable within each species.
* Create boxplot for hindfoot_length. Overlay the boxplot layer on a jitter layer to show actual measurements.

```{r}
ggplot(data = surveys_complete, mapping = aes(x = species_id, y = hindfoot_length)) + 
  geom_jitter(alpha = 0.3, color = "violet") + 
  geom_boxplot(alpha = 0)
```

* Add color to the data points on your boxplot according to the plot from which the sample was taken (plot_id).

```{r}
ggplot(data = surveys_complete, mapping = aes(x = species_id, y = hindfoot_length)) + 
  geom_jitter(aes(color = plot_id)) + 
  geom_boxplot(alpha = 0)
```

* Change the class of plot_id from integer to factor
```{r}
surveys_complete$plot_id <- as.factor(surveys_complete$plot_id)
```

```{r}
ggplot(data = surveys_complete, mapping = aes(x = species_id, y = hindfoot_length)) + 
  geom_jitter(aes(color = plot_id)) + 
  geom_boxplot(alpha = 0)
```

# Plotting time series data

```{r}
yearly_counts <- surveys_complete %>%
  count(year, genus)
```

```{r}
ggplot(data = yearly_counts, aes(x = year, y = n)) +
     geom_line()
```

```{r}
ggplot(data = yearly_counts, aes(x = year, y = n, group = genus)) +
    geom_line()
```

```{r}
ggplot(data = yearly_counts, aes(x = year, y = n, color = genus)) +
    geom_line()
```

# Integrating the pipe operator with ggplot2
  We can also use the pipe operator to pass the data argument to the ggplot() function. The hard part is to remember that to build your ggplot, you need to use + and not %>%.
```{r}
yearly_counts %>% 
    ggplot(mapping = aes(x = year, y = n, color = genus)) +
    geom_line()
```

```{r}
yearly_counts_graph <- surveys_complete %>%
    count(year, genus) %>% 
    ggplot(mapping = aes(x = year, y = n, color = genus)) +
    geom_line()

yearly_counts_graph
```

# Faceting
```{r}
ggplot(data = yearly_counts, aes(x = year, y = n)) +
    geom_line() +
    facet_wrap(facets = vars(genus))
```

```{r}
 yearly_sex_counts <- surveys_complete %>%
                      count(year, genus, sex)
```

```{r}
ggplot(data = yearly_sex_counts, mapping = aes(x = year, y = n, color = sex)) +
  geom_line() +
  facet_wrap(facets =  vars(genus))
```

```{r}
ggplot(data = yearly_sex_counts, 
       mapping = aes(x = year, y = n, color = sex)) +
  geom_line() +
  facet_grid(rows = vars(sex), cols =  vars(genus))
```

```{r}
# One column, facet by rows
ggplot(data = yearly_sex_counts, 
       mapping = aes(x = year, y = n, color = sex)) +
  geom_line() +
  facet_grid(rows = vars(genus))
```

```{r}
# One row, facet by column
ggplot(data = yearly_sex_counts, 
       mapping = aes(x = year, y = n, color = sex)) +
  geom_line() +
  facet_grid(cols = vars(genus))
```
# ggplot2 themes

```{r}
 ggplot(data = yearly_sex_counts, 
        mapping = aes(x = year, y = n, color = sex)) +
     geom_line() +
     facet_wrap(vars(genus)) +
     theme_bw()
```

  Other themes can be 'theme_minimal()','theme_light()', and 'theme_void()'.

## Challenge
  Use what you just learned to create a plot that depicts how the average weight of each species changes through the years.

```{r}
yearly_weight <- surveys_complete %>%
                group_by(year, species_id) %>%
                 summarize(avg_weight = mean(weight))
ggplot(data = yearly_weight, mapping = aes(x=year, y=avg_weight)) +
   geom_line() +
   facet_wrap(vars(species_id)) +
   theme_bw()
```
# Customization

```{r}
ggplot(data = yearly_sex_counts, aes(x = year, y = n, color = sex)) +
    geom_line() +
    facet_wrap(vars(genus)) +
    labs(title = "Observed genera through time",
         x = "Year of observation",
         y = "Number of individuals") +
    theme_bw()
```
```{r}
ggplot(data = yearly_sex_counts, mapping = aes(x = year, y = n, color = sex)) +
    geom_line() +
    facet_wrap(vars(genus)) +
    labs(title = "Observed genera through time",
        x = "Year of observation",
        y = "Number of individuals") +
    theme_bw() +
    theme(text=element_text(size = 16))
```
* Using extrafont package

```{r}
ggplot(data = yearly_sex_counts, mapping = aes(x = year, y = n, color = sex)) +
    geom_line() +
    facet_wrap(vars(genus)) +
    labs(title = "Observed genera through time",
        x = "Year of observation",
        y = "Number of individuals") +
    theme_bw() +
    theme(axis.text.x = element_text(colour = "grey20", size = 12, angle = 90, hjust = 0.5, vjust = 0.5),
                        axis.text.y = element_text(colour = "grey20", size = 12),
                        strip.text = element_text(face = "italic"),
                        text = element_text(size = 16))
```

```{r}
grey_theme <- theme(axis.text.x = element_text(colour="grey20", size = 12, 
                                               angle = 90, hjust = 0.5, vjust = 0.5),
                    axis.text.y = element_text(colour = "grey20", size = 12),
                    text=element_text(size = 16))

ggplot(surveys_complete, aes(x = species_id, y = hindfoot_length)) +
    geom_boxplot() +
    grey_theme
```

## Challenge
  With all of this information in hand, please take another five minutes to either improve one of the plots generated in this exercise or create a beautiful graph of your own. Use the RStudio ggplot2 cheat sheet for inspiration.

Here are some ideas:

* See if you can change the thickness of the lines.
* Can you find a way to change the name of the legend? What about its labels?
* Try using a different color palette (see http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/).

```{r}
grey_theme <- theme(axis.text.x = element_text(colour="grey20", size = 12, 
                                               angle = 90, hjust = 0.5, vjust = 0.5),
                    axis.text.y = element_text(colour = "grey20", size = 12),
                    text=element_text(size = 16))

ggplot(surveys_complete, aes(x = species_id, y = hindfoot_length, color = species_id)) +
    geom_boxplot(size = 1) +
    labs(x = "Hindfoot Length", y = "Species ID") + 
           grey_theme
```

