---
title: "lab5_Hyejoo"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

## US confirmed case bar plot : June VS September
```{r}
Confirmed_State_6_13 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/06-13-2020.csv")) %>%
  filter (Country_Region == "US") %>% 
  group_by(Province_State, Country_Region) %>% 
  summarise(Confirmed = sum(Confirmed)) 
```
```{r}
Confirmed_State_9_13 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/09-13-2020.csv")) %>% 
  filter (Country_Region == "US") %>% 
  group_by(Province_State, Country_Region) %>% 
  summarise(Confirmed = sum(Confirmed)) 
```

```{r}
Confirmed_State_6_13_9_13_joined <- full_join(Confirmed_State_6_13,
      Confirmed_State_9_13, by = c("Province_State")) %>% 
      rename(Confirmed_6_13_2020 = "Confirmed.x", Confirmed_9_13_2020 = "Confirmed.y") %>% 
      select(-Country_Region.x, -Country_Region.y)
head(Confirmed_State_6_13_9_13_joined)
```

```{r}
which(is.na(Confirmed_State_6_13_9_13_joined))
```

```{r, fig.width = 7, fig.height = 10}
Confirmed_State_6_13_9_13_joined_long <- Confirmed_State_6_13_9_13_joined %>% 
  pivot_longer(-c(Province_State),
               names_to = "Date", values_to = "Confirmed")
ggplot(Confirmed_State_6_13_9_13_joined_long, aes(x = Confirmed, y = Province_State)) + 
  geom_bar(stat = 'identity', color = "black", aes(fill = Date))+ 
  labs(title = "US COVID-19 Confirmed Cases",
       subtitle = "June VS September",
       x = "Confirmed Cases", 
       y = "State") +  
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5))
```

## Worldwide Deaths per day

```{r, eval=FALSE}
download.file(url="https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv", 
               destfile = "data/time_series_covid19_deaths_global.csv")
```

```{r}
time_series_deaths <- read_csv("data/time_series_covid19_deaths_global.csv")%>%
  rename(Province_State = "Province/State", Country_Region = "Country/Region")

time_series_deaths_long <- time_series_deaths %>% 
  pivot_longer(-c(Province_State, Country_Region, Lat, Long), names_to = "Date", values_to = "Deaths")
```

```{r}
library(lubridate)
```

```{r}
time_series_deaths_long$Date <- mdy(time_series_deaths_long$Date)
```

```{r}
deaths_worldwide <- time_series_deaths_long %>% 
  group_by(Date) %>% 
  summarise_at(c("Deaths"), sum)
ggplot(deaths_worldwide, aes(Date, Deaths, group = 1)) + 
  geom_point() + 
  geom_line() + 
  labs(x = "Date", y = "Deaths", title = "COVID-19 Deaths Worldwide")
```

## Death per Confirmed cases US
```{r, eval=FALSE}
download.file(url="https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv", 
               destfile = "data/time_series_covid19_confirmed_global.csv")
```

```{r}
time_series_confirmed <- read_csv("data/time_series_covid19_confirmed_global.csv")%>%
  rename(Province_State = "Province/State", Country_Region = "Country/Region")

time_series_confirmed_long <- time_series_confirmed %>% 
  pivot_longer(-c(Province_State, Country_Region, Lat, Long), names_to = "Date", values_to = "Confirmed")

time_series_confirmed_long <- time_series_confirmed_long %>% 
  unite(Key, Province_State, Country_Region, Date, sep = ".", remove = FALSE)
head(time_series_confirmed_long)
```

```{r}
time_series_deaths_long <- time_series_deaths_long %>% 
  unite(Key, Province_State, Country_Region, Date, sep = ".") %>% 
  select(Key, Deaths)
```
 ##### Due to 'reached getOption("max.print")' error (or other undue error), NA couldn't be replaced, thus I made separate time_series_long_joined.csv file. [Here](long-joined.html) is the way how I wrote the time_series_long_joined.csv. 
```{r}
time_series_long_joined <- full_join(time_series_confirmed_long,
              time_series_deaths_long, by = c("Key")) %>% 
              select(-Key)
head(time_series_long_joined)
```
```{r, eval = FALSE}
which(is.na(time_series_long_joined$Deaths))
which(is.na(time_series_long_joined$Confirmed))
```
```{r}
time_series_long_joined <- read_csv("data/time_series_long_joined.csv")
```


```{r}
time_series_long_joined_plus <- time_series_long_joined %>% 
  mutate(Deaths_per_Confirmed = Deaths/Confirmed)
time_series_long_joined_plus$Deaths_per_Confirmed[is.nan(time_series_long_joined_plus$Deaths_per_Confirmed)] <- 0
```

```{r}
time_series_long_joined_counts <- time_series_long_joined_plus %>% 
  select( , -c(3,4,7,8)) %>%
  pivot_longer(-c(Province_State, Country_Region, Date),
               names_to = "Report_Type", values_to = "Counts")
head(time_series_long_joined_counts)
```

```{r}
time_series_long_joined_plus$Date <- mdy(time_series_long_joined_plus$Date)
```

```{r}
time_series_long_joined_plus %>% 
  group_by(Country_Region,Date) %>% 
  summarise_at(c("Confirmed", "Deaths", "Deaths_per_Confirmed"), sum) %>% 
  filter (Country_Region == "US") %>% 
    ggplot(aes(x = Date,  y = Deaths_per_Confirmed)) + 
    geom_point() +
    geom_line() + 
    ggtitle("US COVID-19 Deaths per Confirmed Cases")
```

## Top 10 Countires for Deaths

```{r}
library(tidyverse)
library(DT)
```


```{r}
time_series_top_10 <- time_series_long_joined_plus %>% 
  group_by(Country_Region) %>% 
  summarize(Confirmed = sum(Confirmed),
            Deaths = sum(Deaths),
            Deaths_per_Confirmed = sum(Deaths_per_Confirmed))%>% 
  arrange(desc(Deaths)) %>% 
  slice(1:10)
datatable(time_series_top_10)
```

```{r}
time_series_long_joined_plus %>% 
  group_by(Country_Region,Date) %>% 
  summarise_at(c("Confirmed", "Deaths", "Deaths_per_Confirmed"), sum) %>% 
  filter(Country_Region %in% c("US", "Brazil", "United Kingdom", "Italy", "Mexico", "India", "France", "Spain", "Iran", "Peru")) %>% 
  ggplot(aes(x = Date,  y = Deaths, color = Country_Region)) + 
  geom_point() + 
  geom_line() + 
  ggtitle("Top 10 COVID-19 Death Countries") + 
  theme(text=element_text(size = 16))
```
## Facet_Wraps for Top 10 COVID-19 Deaths

```{r}
library(scales)
```

```{r}
Thousand <- function (x) { number_format(accuracy = 1,
                                   scale = 1/1000,
                                   suffix = "k",
                                   big.mark = ",")(x) }
```

```{r, fig.width = 7, fig.height = 7}
time_series_long_joined_plus %>% 
  group_by(Country_Region,Date) %>% 
  summarise_at(c("Confirmed", "Deaths", "Deaths_per_Confirmed"), sum) %>% 
  filter(Country_Region %in% c("US", "Brazil", "United Kingdom", "Italy", "Mexico", "India", "France", "Spain", "Iran", "Peru")) %>% 
  ggplot(aes(x = Date,  y = Deaths, color = Country_Region)) + 
  geom_point() + 
  geom_line() + 
  ggtitle("Top 10 COVID-19 Death Countries") + 
  facet_wrap(~Country_Region, ncol=3, scales="free_y") + 
  scale_y_continuous(label = Thousand) + 
  theme_linedraw() + 
  theme(strip.text = element_text(face = "italic"), text=element_text(size = 16), legend.position="none")
```

## Facet_Wraps for Confirmed Cases for All US States and Territories

```{r, eval = FALSE}
download.file(url="https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_US.csv", 
               destfile = "data/time_series_covid19_confirmed_US.csv")
```

```{r}
time_series_us_confirmed <- read_csv("data/time_series_covid19_confirmed_US.csv")
head(time_series_us_confirmed)
```

```{r}
time_seriese_us_confirmed_long <- time_series_us_confirmed %>% 
  select( , -c(1:6)) %>% 
  pivot_longer(-c(Province_State, Country_Region, Lat, Long_, Combined_Key), names_to = "Date", values_to = "Confirmed")
which(is.na(time_seriese_us_confirmed_long$Confirmed))
head(time_seriese_us_confirmed_long)
```
```{r}
time_seriese_us_confirmed_long$Date <- mdy(time_seriese_us_confirmed_long$Date)
```

```{r, fig.width = 12, fig.height = 12}
time_seriese_us_confirmed_long %>% 
  group_by(Province_State,Date) %>% 
  summarise_at(c("Confirmed"), sum) %>% 
  ggplot(aes(Date, Confirmed, group = 1)) + 
  geom_line() + 
  ggtitle("US COVID-19 Confirmed Cases") + 
  facet_wrap(~Province_State, ncol=6, scales="free_y") + 
  scale_y_continuous(label = Thousand) + 
  theme_bw() + 
  theme(axis.text.x = element_text(size = 10), axis.text.y = element_text(size = 10),strip.text = element_text(size = 10, face = "italic"), text=element_text(size = 16))
```
