---
title: "Lab6: Data Maps and Interactive Graphs from the COVID-19 reporting data"
date: September 28, 2020
output:
  rmarkdown::html_document:
    theme: lumen
bibliography: data/covid_references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
library(maps)
library(mapdata)
library(lubridate)
library(viridis)
library(wesanderson)
library(RColorBrewer)
library(plotly)
```

# Excercise 1
  For “COVID-19 Deaths” graph, summarize the counts for each Country on the graph and update the graph to 9/26/2020.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
report_9_26_2020 <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/09-26-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  filter(Long != "NA") %>% 
  filter(Lat != "NA") %>% 
  group_by(Country_Region) %>% 
  summarize(Confirmed = sum(Confirmed), Deaths = sum(Deaths), Lat= median(Lat), Long= median(Long))

# Read in the world map data
world <- as_tibble(map_data("world"))

# Change the name of some countries
world <- as_tibble(map_data("world")) %>% 
 mutate(region = str_replace_all(region, c("USA" = "US", "Czech Republic" = "Czechia",  
        "Ivory Coast" = "Cote d'Ivoire", "Democratic Republic of the Congo" = "Congo (Kinshasa)", 
        "Republic of Congo" = "Congo (Brazzaville)", "South Korea" = "Korea, South", "Swaziland" = 	
"Switzerland", "Taiwan" = "Taiwan*", "UK" = "United Kingdom", "Myanmar" = 	
"Burma")))

# Join the covid report with the map data
country_join <- left_join(world, report_9_26_2020, by = c("region" = "Country_Region")) %>% 
  filter(Lat != "NA") %>% 
  filter(Long != "NA")

# Create the graph
ggplotly(
ggplot(world, mapping = aes(x = long, y = lat, text = region, group = group)) + 
  coord_fixed(1.3) + 
  geom_polygon(data = country_join, aes(fill = Deaths), color = "black") +
  scale_fill_gradientn(colours = 
                         wes_palette("Zissou1", 100, type = "continuous")) +
  labs(title = "COVID-19 Deaths"))
```


```{r}
# Create the graph
ggplotly(
ggplot(world, mapping = aes(x = long, y = lat, text = region, group = group)) + 
  coord_fixed(1.3) + 
  geom_polygon(data = country_join, aes(fill = Confirmed), color = "black") +
  scale_fill_gradientn(colours = 
                         wes_palette("Zissou1", 100, type = "continuous")) +
  labs(title = "COVID-19 Confirmed Cases"))
```



# Excercise 2
  Update Anisa Dhana’s graph layout of the US to 9/26/2020.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
report_09_26_2020 <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/09-26-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  filter(Country_Region == "US") %>% 
  filter (!Province_State %in% c("Alaska","Hawaii", "American Samoa",
                  "Puerto Rico","Northern Mariana Islands", 
                  "Virgin Islands", "Recovered", "Guam", "Grand Princess",
                  "District of Columbia", "Diamond Princess")) %>% 
  filter(Lat > 0)

mybreaks <- c(1, 100, 1000, 10000, 10000)

ggplot(report_09_26_2020, aes(x = Long, y = Lat, size = Confirmed)) +
    borders("state", colour = "white", fill = "grey90") +
    geom_point(aes(x=Long, y=Lat, size=Confirmed, color=Confirmed),stroke=F, alpha=0.7) +
    scale_size_continuous(name="Cases", range=c(1,7), breaks=mybreaks, labels = c("1-99", "100-999", "1,000-9,999", "10,000-99,999", "100,000+")) +
    scale_color_viridis_c(option="magma",name="Cases",breaks=mybreaks, trans="log", labels = c("1-99","100-999", "1,000-9,999", "10,000-99,999", "100,000+")) +
    theme_void() + 
    guides( colour = guide_legend()) +
    labs(title = "Anisa Dhana's lagout for COVID-19 Confirmed Cases in the US'") +
    theme(
      legend.position = "bottom",
      text = element_text(color = "#22211d"),
      plot.background = element_rect(fill = "#ffffff", color = NA), 
      panel.background = element_rect(fill = "#ffffff", color = NA), 
      legend.background = element_rect(fill = "#ffffff", color = NA)
    ) +
    coord_fixed(ratio=1.5)
```

# Exercise 3
  Update the above graph “Number of Confirmed Cases by US County” to 9/26/2020 and use a different color scheme or theme

```{r, echo=FALSE, message=FALSE, warning=FALSE}
report_09_26_2020 <-  read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/09-26-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  unite(Key, Admin2, Province_State, sep = ".") %>% 
  group_by(Key) %>% 
  summarize(Confirmed = sum(Confirmed)) %>% 
  mutate(Key = tolower(Key))

# Get and format map data
us <- map_data("state")
counties <- map_data("county")%>% 
  unite(Key, subregion, region, sep = ".", remove = FALSE)
state_join <- left_join(counties, report_09_26_2020, by = c("Key"))

# Plot map
ggplot(us, aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
  borders("state", colour = "black") +
  geom_polygon(data = state_join, aes(fill = Confirmed)) +
  scale_fill_gradientn(colors = brewer.pal(n = 5, name = "BuPu"),trans = "log10", breaks = c(1, 10, 100, 1000, 10000, 100000), na.value = "White") +
  ggtitle("Number of Confirmed Cases by US County") +
  theme_bw()
```

# Exercise 4
  Make an interactive plot with data of Minnesota.
```{r, echo=FALSE, message=FALSE, warning=FALSE}
daily_report <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/09-26-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  filter(Province_State == "Minnesota") %>% 
  group_by(Admin2) %>% 
  summarize(Confirmed = sum(Confirmed)) %>% 
  mutate(Admin2 = tolower(Admin2))

# Get and format state map data
us <- map_data("state")
mn_us <- subset(us, region == "minnesota")
counties <- map_data("county")
mn_county <- subset(counties, region == "minnesota")
state_join <- left_join(mn_county, daily_report, by = c("subregion" = "Admin2")) 

# Plot state map
ggplotly(ggplot(data = mn_county, mapping = aes(x = long, y = lat, group = group)) + 
           coord_fixed(1.3) + 
           geom_polygon(data = state_join, aes(fill = Confirmed), colour = alpha("grey", 1 / 2)) + 
          scale_fill_viridis_c(option = 'magma', direction = -1) +
  ggtitle("COVID-19 Confirmed Cases in Minnesota") + 
    labs(x=NULL, y=NULL)+ theme_bw())
```

# Exercise 5
  Create a report with static maps and interactive graphs that is meant to be read by others.
  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
jpreport_09_26_2020 <-  read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/09-26-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  rename(region = "Province_State") %>% 
  filter(Country_Region == "Japan") %>%
  group_by(region) %>% 
  summarize(Confirmed = sum(Confirmed), Deaths = sum(Deaths))

# Get and format map data
jp <- map_data("japan")
jpstate_join <- left_join(jp, jpreport_09_26_2020, by = "region")

# Plot map
ggplot(jp, aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) +
  geom_polygon(data = jpstate_join, aes(fill = Confirmed), color = "White" ) +
  scale_fill_gradientn(colors = brewer.pal(n = 10, name = "BuPu"),trans = "log", breaks = c(1, 100, 250, 500, 1000, 2000), na.value = "White", guide = guide_legend(keyheight = unit(3, units = "mm"), keywidth=unit(12, units = "mm"), label.position = "bottom", title.position = 'top', nrow=1)) +
  ggtitle("COVID-19 Confirmed Cases in Japan") +
  theme_void()+
  theme(
    plot.title = element_text(size= 14, hjust=0.01, margin = margin(b = -0.1, t = 0.4, l = 2, unit = "cm")),
    legend.position = c(0.7, 0.09)
  ) 
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Plot state map
ggplotly(ggplot(jp, mapping = aes(x = long, y = lat, group = group)) + 
           coord_fixed(1.3) + 
           geom_polygon(data = jpstate_join, aes(fill = Deaths), colour = alpha("black", 1/3))  +
         scale_fill_viridis_c(option = 'inferno', direction = -1)+
  ggtitle("COVID-19 Deaths in Japan (Interactive Graph)") + 
    labs(x=NULL, y=NULL) +
  theme(panel.border = element_blank()) +
  theme(panel.background = element_blank()) +
  theme(axis.ticks = element_blank()) +
  theme(axis.text = element_blank()))
```

All data above from "JHU CSSE COVID-19 Data" [@JHU] and Dong et al. 2020 [@DongE].

# REFERENCES

