---
title: "lab9x Programming in R - Control Structures and Functions
"
output: html_document
---

## Loops
* if and else: testing a condition and acting on it
* for: execute a loop a fixed number of times
* while: execute a loop while a condition is true

## Important Operators
### Comparison operators
* == (equal)
* != (not equal)
* > (greater than)
* >= (greater than or equal)
* < (less than)
* <= (less than or equal)

### Logical operators
* & (and)
* | (or)
* ! (not)

## Example of if-else statement

```{r}
x <- runif(1, 0, 100)  
if(x > 5) {
        y <- 100
} else {
        y <- 0
}
```

```{r}
x <- 10
if (x < 0) {
print("-")
} else if (x > 0) {
print("+")
} else
print("Zero")
```

## for Loops
```{r}
for(i in 1:5) {
        print(i)
}
```

```{r}
x <- c("x", "b", "r", "l", "e")
for(i in 2:4) {
        ## Print out each element of 'x'
        print(x[i])  
}
```

```{r}
## Generate a sequence based on length of 'x'
for(i in seq_along(x)) {   
        print(x[i])
}
```

```{r}
for(letter in x) {
        print(letter)
}

```

```{r}
## One line loops
for(i in 2:5) print(x[i])
```

## Nested for loops

```{r}
x <- matrix(1:6, 2, 3)
for(i in seq_len(nrow(x))) {
        for(j in seq_len(ncol(x))) {
                print(x[i, j])
        }   
}
```



## while loops
```{r}
count <- 0
while(count > -10) {
        print(count)
        count <- count - 2
}
```

```{r}
z <- 5
set.seed(1)
while(z >= 3 && z <= 10) {
        coin <- rbinom(1, 1, 0.5)
        
        if(coin == 1) {  ## random walk
                z <- z + 1
        } else {
                z <- z - 1
        } 
}
print(z)
```

## Functions

* Functions can be passed as arguments to other functions. This is very handy for the various apply functions, like lapply() and sapply() that we will learn below.

* Functions can be nested, so that you can define a function inside of another function

```{r}
f <- function() {
        ## This is an empty function
}
## Functions have their own class
class(f)  
```

```{r}
## Execute this function
f()
```

```{r}
f <- function() {
        cat("Hello, world!\n")
}
f()
```

```{r}
f <- function(num) {
        for(i in seq_len(num)) {
                cat("Hello, world!\n")
        }
}
f(4)
```

```{r}
f <- function(num) {
        hello <- "Hello, world!\n"
        for(i in seq_len(num)) {
                cat(hello)
        }
        chars <- nchar(hello) * num
        chars
}
meaningoflife <- f(4)
```

```{r}
print(meaningoflife)
```

```{r}
f <- function(num = 1) {
        hello <- "Hello, world!\n"
        for(i in seq_len(num)) {
                cat(hello)
        }
        chars <- nchar(hello) * num
        chars
}
f()    ## Use default value for 'num'
```

```{r}
f(3)   ## Use user-specified value
```

```{r}
f(num = 3)
```

